Disturbance-aware minimum-time planning strategies for motorsport vehicles with probabilistic safety certificates

### 2025-10-01 intro.tex
[MGa] mods to comments to Cole

Despite these advances, most planning frameworks still fail to incorporate robustness in the planning phase itself. A notable exception is~\cite{Timings:RobustLaptimeSimulation:2014a}, which first plans a nominal trajectory via MPC and then employs a feedback MPC to counteract disturbance realizations inferred from road-surface roughness via a ride model. An inspiring aspect is the explicit trade-off between drivability and the control effort required to stay close to the nominal path. While there are many similarities with our setting, the key differences are methodological: their approach uses a two-level, MPC-based scheme with discrete disturbance realizations in the robust layer, whereas our lap-time simulation/planning is posed as a single large-scale NLP in which robustness to (Gaussian) disturbances is embedded directly within one optimization problem.

>>>>

Despite these advances, most planning frameworks still fail to incorporate robustness in the planning phase itself. A notable exception is~\cite{Timings:RobustLaptimeSimulation:2014a}, which first plans a nominal trajectory via MPC and then employs a feedback MPC to counteract disturbance realizations inferred from road-surface roughness via a ride model. An inspiring aspect is the trade-off between vehicle/driver performance and the control effort demanded of the driver to stay close to the nominal path despite disturbances, which essentially defines \emph{drivability}. While there are many similarities with our setting, the key differences are methodological: their approach uses a two-level, MPC-based scheme with discrete disturbance realizations in the robust layer, whereas our lap-time simulation/planning is posed as a single large-scale NLP in which robustness to (Gaussian) disturbances is embedded directly within one optimization problem.
